Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy

### 2020-10-07 manuscript.Rmd
fixing some typos

Their final sample of participants, after data exclusion, included 26 participants (data available at https://osf.io/hj7tz/). The EMG data is depicted in Figure \@ref(fig:general) by condition (where `BAS`, `DIS`, and `RUM` refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (frontalis, `FRO`; orbicularis oris inferior `OOI`; and oribicularis oris superior, `OOS`). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.44$, meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually $0.36$).

@@

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis of d = 0), an effect size of d = 0.36 [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of d = 0.72 [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis (although the "true" effect size is d = 0). When the "true" effect size is of d = 0.36, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of d = 0.72, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.^[To assess the extent to which the BF10 computed for the OOI by @moffatt_inner_2020 is "surprising" given or "compatible" with the sampling distribution of BFs under the hypothesis of an effect size of d = 0.36, we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to `r round(1 - mean(results$BF10 > 0.278), 2)`.]

>>>>

Their final sample of participants, after data exclusion, included 26 participants (data available at https://osf.io/hj7tz/). The EMG data is depicted in Figure \@ref(fig:general) by condition (where BAS, DIS, and RUM refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (frontalis, FRO; orbicularis oris inferior, OOI; and orbicularis oris superior, OOS). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

We suggest the (a priori) power of the study ran by @moffatt_inner_2020 was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter [which seems reasonable given the high inter-individual variability in such effects, cf. the next section but also @nalborczyk_can_2020], therefore the a priori power of the main statistical test from @moffatt_inner_2020 was around $0.42$, meaning that they had less than 1 chance out of 2 to find a significant effect (given that the population effect size was actually $0.36$).

@@

First, we simulated 10.000 datasets (for $N = 26$) under the assumption of either no effect (i.e., the null hypothesis of $d = 0$), an effect size of $d = 0.36$ [i.e., the supposed target effect size in @moffatt_inner_2020], or an effect size of $d = 0.72$ [i.e., the effect size reported in @nalborczyk_orofacial_2017]. As shown in Figure \@ref(fig:bf-dance), the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, `r round(prop_d0 * 100, 2)`% of the computed BFs are above 0 and hence support the alternative hypothesis (although the "true" effect size is $d = 0$). When the "true" effect size is of $d = 0.36$, `r round(100 - prop_d0.36 * 100, 2)`% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the "true" effect size is of $d = 0.72$, `r round(100 - prop_d0.72 * 100, 2)`% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.^[To assess the extent to which the $\text{BF}_{10}$ computed for the OOI by @moffatt_inner_2020 is "surprising" given or "compatible" with the hypothesis of an effect size of $d = 0.36$, we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to `r round(1 - mean(results$BF10 > 0.278), 2)`.]

### 2020-10-07 manuscript.tex
fixing some typos

Wordcount (excluding abstract, references, tables, and figures): 4536

@@

Their final sample of participants, after data exclusion, included 26 participants (data available at \url{https://osf.io/hj7tz/}). The EMG data is depicted in Figure \ref{fig:general} by condition (where \texttt{BAS}, \texttt{DIS}, and \texttt{RUM} refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (frontalis, \texttt{FRO}; orbicularis oris inferior \texttt{OOI}; and oribicularis oris superior, \texttt{OOS}). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.44\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the effect in the population was actually \(0.36\)).

@@

First, we simulated 10.000 datasets (for N = 26) under the assumption of either no effect (i.e., the null hypothesis of d = 0), an effect size of d = 0.36 (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of d = 0.72 (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}). As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, 6.96\% of the computed BFs are above 0 and hence support the alternative hypothesis (although the ``true'' effect size is d = 0). When the ``true'' effect size is of d = 0.36, 71.83\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of d = 0.72, 25.58\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.\footnote{To assess the extent to which the BF10 computed for the OOI by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) is ``surprising'' given or ``compatible'' with the sampling distribution of BFs under the hypothesis of an effect size of d = 0.36, we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to 0.29.}

>>>>

Wordcount (excluding abstract, references, tables, and figures): 4549

@@

Their final sample of participants, after data exclusion, included 26 participants (data available at \url{https://osf.io/hj7tz/}). The EMG data is depicted in Figure \ref{fig:general} by condition (where BAS, DIS, and RUM refer to the baseline, distraction, and rumination conditions, respectively) and by muscle (frontalis, FRO; orbicularis oris inferior, OOI; and orbicularis oris superior, OOS). This figure shows that the average natural logarithm of the EMG peak amplitude recorded over the FRO was at similar levels in the baseline and distraction conditions, but was much higher in the rumination condition. However, the average natural logarithm of the EMG peak amplitude recorded over the OOI and OOS muscles was higher than baseline in both the rumination and distraction conditions, with a slight increase from distraction to rumination (both on the mean and median).

@@

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.42\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the population effect size was actually \(0.36\)).

@@

First, we simulated 10.000 datasets (for \(N = 26\)) under the assumption of either no effect (i.e., the null hypothesis of \(d = 0\)), an effect size of \(d = 0.36\) (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of \(d = 0.72\) (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}). As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, 6.96\% of the computed BFs are above 0 and hence support the alternative hypothesis (although the ``true'' effect size is \(d = 0\)). When the ``true'' effect size is of \(d = 0.36\), 71.83\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of \(d = 0.72\), 25.58\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.\footnote{To assess the extent to which the \(\text{BF}_{10}\) computed for the OOI by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) is ``surprising'' given or ``compatible'' with the hypothesis of an effect size of \(d = 0.36\), we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to 0.29.}
